---
title: Lab01 Data Types, Especially Lists in R
author:
- name: Goran S. Milovanovic, PhD
abstract: 
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: true
    theme: cosmo
    highlight: textmate
---

***
# Labs 01: A Very Basic Introduction to Data Types and Lists in R 
**Feedback** should be send to `goran.milovanovic@datakolektiv.com`. 
### 0. Data Types in R

Some datasets come with the R programming language, e.g., the famous `iris`:

```{r echo = T, eval = T, message = F}
data(iris)
head(iris, 10)
```

The function `head()`: show me the first n rows of a data.frame! Similarly, the function `tail()`:

```{r echo = T, eval = T, message = F}
tail(iris, 10)
```

returns the last n rows of a data.frame.

```{r echo = T, eval = T, message = F}
my_vector <- c(1, 7, 9, 10, 14, 22, 3.14, 2.71, 99)
head(my_vector, 5)
```

```{r echo = T, eval = T, message = F}
tail(my_vector, 5)
```

### 1. Lists

*Lists* are very important in R. Let's create one:

```{r echo = T, eval = T, message = F}
my_list <- list(element_1 = 1,
                element_2 = "Belgrade", 
                element_3 = TRUE)
str(my_list)
```

The function `str()` in R is generic, meaning it can be used on objects of *different classes*. We will discuss this further during the course.

Here’s another list describing a person:

```{r echo = T, eval = T, message = F}
person <- list(name = "Mark",
               family_name = "Smith",
               phone = "+381661722838383", 
               email = "mark.smith@rcourses.org", 
               age = 40,
               gender = "M", 
               employed = TRUE)
person
```

Anything can be an element of a list in R—even an entire data.frame:

```{r echo = T, eval = T, message = F}
person <- list(name = "Mark",
               family_name = "Smith",
               phone = "+381661722838383", 
               email = "mark.smith@rcourses.org", 
               age = 40,
               gender = "M", 
               employed = TRUE, 
               favorite_dataset = "iris", 
               favorite_dataset_source = iris)
```

Lists can be nested:

```{r echo = T, eval = T, message = F}
ll <- list(e1 = 10, 
           e2 = 20, 
           e3 = list(
             e1 = 20,
             e2 = 40,
             e3 = 15
           ), 
           e4 = 40,
           e5 = list(
             e1 = 12
           ))
ll
```

For example, data structures describing people via R lists:

```{r echo = T, eval = T, message = F}
persons <- list(name = c("Mark", "Jane"),
                family_name = c("Smith", "Doe"),
                phone = c("+381661722838383", "+381661722838384"),
                email = c("mark.smith@rcourses.org", "jane.doe@rcourses.org"), 
                age = c(40, 42),
                gender = c("M", "F"),
                employed = c(TRUE, FALSE)
                )
```

Accessing list elements:

```{r echo = T, eval = T, message = F}
persons[[1]]
```

Accessing elements of a *named* list:

```{r echo = T, eval = T, message = F}
persons$family_name[2]
```

The logic of structuring data is *key* in Data Science. Here’s a better way to describe people, e.g., employees in a company, using lists:

```{r echo = T, eval = T, message = F}
persons <- list(
  p1 = list(name = "Mark",
            family_name = "Smith",
            phone = "+381661722838383",
            email = "mark.smith@rcourses.org",
            age = 40,
            gender = "M",
            employed = TRUE
            ),
  p2 = list(name = "Jane",
            family_name = "Doe",
            phone = "+381661722838385",
            email = "jane.doe@rcourses.org",
            age = 42,
            gender = "F",
            employed = FALSE
            )
)
```

Accessing list elements:

```{r echo = T, eval = T, message = F}
persons[[1]]
```

### 2. ```data.frame``` Class

This class is essentially the central component we work with in the R programming language:

```{r echo = T, eval = T, message = F}
str(iris)
```

Unique elements of the variable (column, field, however you prefer) `Species` in `iris` using the function `unique()`:

```{r echo = T, eval = T, message = F}
unique(iris$Species)
```

Nesting function calls in R—something we do quite often. For example, "give me the length (length) of a vector obtained by listing the unique elements (unique) of the Sepal.Length column in iris" is written in R as:

```{r echo = T, eval = T, message = F}
length(
  unique(
    iris$Sepal.Length
    )
  )
```

`length()` returns the length of a vector or list:

```{r echo = T, eval = T, message = F}
length(iris$Sepal.Length)
```

`dim()` gives the dimensions of a data.frame, for example:

```{r echo = T, eval = T, message = F}
dim(iris)
```

`dim()` returns a vector (e.g., the number of rows and columns for the data.frame class). Remember that the result of a function in R can also be "subsetted," i.e., you can extract only the part of the result you need by indexing. For example, how many rows does `iris` have:

```{r echo = T, eval = T, message = F}
dim(iris)[1]
```

***
License: [GPLv3](http://www.gnu.org/licenses/gpl-3.0.txt)
This Notebook is free software: you can redistribute it and/or modify it under the terms of the GNU General Public License as published by the Free Software Foundation, either version 3 of the License, or (at your option) any later version.
This Notebook is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License for more details.
You should have received a copy of the GNU General Public License along with this Notebook. If not, see <http://www.gnu.org/licenses/>.
