---
title: Notes for Session04 More Functional Programming and Vectorization
author:
- name: Goran S. Milovanovic, PhD
abstract: 
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: true
    theme: cosmo
    highlight: textmate
---

***
# Notes 04: More Functional Programming and Vectorization

[Advanced R 2nd edition, Hadley Wickham, Chapter: Functional programming](https://adv-r.hadley.nz/fp.html)

## Vectorization and Recycling

#### Element-wise operations

```{r echo=TRUE}
### devsify.ai
### Notes for Session04 More Functional Programming
v1 <- c(1,2,3,4,5,17)
print(v1)
```

```{r echo=TRUE}
### devsify.ai
### Notes for Session04 More Functional Programming
a <- 5
print(a[1])
print(a)
```

```{r echo=TRUE}
### devsify.ai
### Notes for Session04 More Functional Programming
v1 <- 5
v2 <- c(2,3,4)
print(v1 + v2)
```

```{r echo=TRUE}
### devsify.ai
### Notes for Session04 More Functional Programming
v1 <- c(1,2,3)
v2 <- c(2,3,4)
print(v1 + v2)
```

```{r echo=TRUE}
### devsify.ai
### Notes for Session04 More Functional Programming
v1 <- c(1,2)
v2 <- c(2,3,4)
print(v1 + v2)
```

```{r echo=TRUE}
### devsify.ai
### Notes for Session04 More Functional Programming
v1 <- 5
v2 <- c(2,3,4)
print(v2 + v1)
```


## Lapply(), Sapply(), Apply()

```{r echo=TRUE}
### devsify.ai
### Notes for Session04 More Functional Programming

v1 <- 1:100

# - power() function
power <- function(x) {
  # code blocks...
  return(x^2)
}

# - silly functional:
squares <- lapply(v1, power)
class(squares)

# - sapply():
squares <- sapply(v1, power)
class(squares)

# - apply
m1 <- matrix(data = c(1,2,3,4),
             ncol = 2)
sum(c(1,2))
apply(X = m1, MARGIN = 1, FUN = sum)
apply(X = m1, MARGIN = 2, FUN = sum)
```


## Reduce()

```{r echo=TRUE}
### devsify.ai
### Notes for Session04 More Functional Programming

```

## Map()

```{r echo=TRUE}
### devsify.ai
### Notes for Session04 More Functional Programming

c1 <- c("Paris", "London")
c2 <- c("France", "UK")
m <- Map(paste, c1, c2)
class(m)
print(m)

# a binary function
f2 <- function(a, b) {
  return(
    a^2 + b
  )
}

v1 <- 1:20
v2 <- 21:40
m <- Map(f2, v1, v2)
print(m)

```

## Function Factories

```{r echo=TRUE}
### devsify.ai
### Notes for Session04 More Functional Programming

```


***
License: [GPLv3](http://www.gnu.org/licenses/gpl-3.0.txt)
This Notebook is free software: you can redistribute it and/or modify it under the terms of the GNU General Public License as published by the Free Software Foundation, either version 3 of the License, or (at your option) any later version.
This Notebook is distributed in the hope that it will be useful, but WITHOUT ANY WARRANTY; without even the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the GNU General Public License for more details.
You should have received a copy of the GNU General Public License along with this Notebook. If not, see <http://www.gnu.org/licenses/>.

